---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Basic setups
```{r setup}
library(epiflow)
```

### Load in packages needed and build-in dataset
Data source: https://hrs.isr.umich.edu/data-products/access-to-public-data (public available)
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
# Load in build-in dataset -- new_hrs.rda
system.file("data", "new_hrs.rda", package = "epiflow")
```

### Look at the structure of the data
```{r message=FALSE, warning=FALSE}
datatable(new_hrs)
```

&nbsp;
&nbsp;

### Variables:
**HHID, PN**: identification number  
**AD_cat**: cognitive status -- Normal, Cognitive Impairment Non Dementia (CIND), Dementia  
**AD_bin**: binary variable -- dementia (1) or not (0)  
**age, sex, education**: self-reported age, sex, years of education  
**HDL, TC**: blood High Density Lipoprotein (HDL), total cholesterol (TC), mg/dL  
**HDL_clinic, TC_clinic**: clinical level of HDL/TC -- at risk low (1) HDL: <40 mg/dL for male and <50 mg/dL for female; at risk high (1) TC: ≥240 mg/dL  
**PGS_HDL, PGS_TC**: polygenic risk score for HDL/TC  

&nbsp;
&nbsp;

# Univariate and bivariate analysis
## 1. Make table one 
There are a lot of existing useful packages for making table one. I prefer to use *compareGroups* package, see details at https://cran.r-project.org/web/packages/compareGroups/index.html.

## 2. make_heatmap( ): Check collinearity (associations across all covariates)
### Inputs:
* __dataset:__ A dataset of interest (should all be numeric variables).  
* **sig_level:** Significant level to be marked in the heat map, 0.05 by default.  
* **lab_size:** Size for label.  
* **fontsize:** Size for font.  
* **width:** Output figure width.  
* **height:** Output figure height.  
* **save_name:** Saved name for your figure.  
* **output_path:** File path for output figure.

### Output:
* A heat map for correlations across all covariates.  
```{r fig.height = 4.5, fig.width = 4.5}
# Make a numerical dataset to plot
hrs_toplot <-
  new_hrs %>% 
  select(AD_bin, age, sex, education, HDL, TC, PGS_HDL, PGS_TC)
# Specify output path
output_path = "/Users/Mingzhou/Desktop/Materials/General R/output/"
# Make a heat map
make_heatmap(hrs_toplot, sig_level = 0.05, lab_size = 3, fontsize = 12, width = 4.5, height = 4.5, 'example_heatmap.png', output_path)
```

&nbsp;
&nbsp;

# Regression analysis
## Logistic regression
### 1. List a couple of models of interest
```{r}
crude <- glm(AD_bin ~ age, data = new_hrs, family = "binomial")
adj_1 <- glm(AD_bin ~ age + sex, data = new_hrs, family = "binomial")
adj_2 <- glm(AD_bin ~ age + sex + education, data = new_hrs, family = "binomial")
```

### 2. extract_OR( ): extract OR for a single model

### Inputs:
* __regress_obj:__ An object given by glm() function
* **index_lst:** A list of index(s) of variable(s) of interest
* **mark:** A mark created by users to identify model(s)

### Output:
* A table of extracted model values

#### For single variable:
```{r message=FALSE, warning=FALSE}
tab_crude <- extract_OR(crude, 2, 'crude_age')
datatable(tab_crude)
```

#### For multiple variables:
```{r message=FALSE, warning=FALSE}
tab_adj2 <- extract_OR(adj_2, c(2, 3, 4), 'adj_full')
datatable(tab_adj2)
```

### 3. make_OR_table( ): make an OR table for all models at once

### Inputs:
* __lst_model:__ A list of logistic regression models
* **index_lst:** A list of index(s) of variable(s) of interest
* **model_head:** The mark created by users to identify model(s)

### Output:
* A table of extracted model values for multiple models

```{r message=FALSE, warning=FALSE}
lst_model <- c('crude', 'adj_1', 'adj_2')
# To look at all values for variabel 'age' at the same time
tab_OR_all <- make_OR_table(lst_model, 2, 'age_')
datatable(tab_OR_all)
```

### 4. Population attributable fraction (PAF) calculation

### Inputs:
* __object:__ A regression object given by glm() function
* __data:__ An optional data frame, list or environment (or object coercible by as.data.frame to a data frame) containing the variables in the model. If not found in data, the variables are taken from environment (formula), typically the environment from which the function is called.
* **exposure:** The name of the exposure variable as a string.
* **mark:** A mark created by the user to identify model(s)
* **case.control:** Can be set to TRUE if the data is from a non-matched case control study. By default case.control is set to FALSE which is used for cross-sectional sampling designs.
* **confidence.level:** User-specified confidence level for the confidence intervals. If not specified it defaults to 95 percent. Should be specified in decimals such as 0.95 for 95 percent.

### Output:
* A table of values for PAF.

```{r}
pop_crude <- glm(AD_bin ~ sex, data = new_hrs, family = 'binomial')
tab_AF <- extract_AF(object = pop_crude, data = new_hrs, exposure = "sex", mark = 'crude_example')
datatable(tab_AF)
```

## Linear regression
Usage of functions are similar to logistic regression:
```{r}
linear_crude <- lm(HDL ~ age, data = new_hrs)
linear_adj_1 <- lm(HDL ~ age + sex, data = new_hrs)
linear_adj_2 <- lm(HDL ~ age + sex + education, data = new_hrs)

lst_model <- c('linear_crude', 'linear_adj_1', 'linear_adj_2')
# To look at all values for variabel 'age' at the same time
tab_linear_all <- make_coeff_table_lm(lst_model, 2, 'age_')
datatable(tab_linear_all)
```

&nbsp;
&nbsp;

# Mendelian randomization
### MR_assumption_test( ): Test for MR assumptions (Strict tests for reference, not recommended in real analysis)

### Inputs:
* __gene_instrument:__ The gene instrument used in the MR analysis (should be single SNP/polygenic score).
* **exposure:** Exposure of the MR model.
* **outcome:** Outcome of the MR model.
* **covariates:** Covariates adjusted in the model (default as NULL).
* **data_name:** Name of the dataset. Please notice that the dataset should have NO missing value.

### Output:
* A set of TRUE/FALSE indicating if the assumptions hold. If the MR assumptions hold, all value should be TRUE.
```{r}
# Drop the missing values
mr_data <-
  new_hrs %>% 
  drop_na()
# Test for MR assumptions
MR_assumption_test(gene_instrument = 'PGS_HDL', exposure = 'HDL_clinic', 
                   outcome = 'AD_bin', covariates = 'age + sex + education', 
                   data_name = 'mr_data')
```

### get_MR_value( ): Get MR results

### Inputs:
* __reg_object1:__ A regression model given by glm() function: exposure ~ gene_instrument (+ covariates)
* **reg_object2:** A regression model given by glm() function: outcome ~ gene_instrument (+ covariates)
* **mark:** Mark of the MR model

### Output:
* A table of extracted results from Mendelian randomization

Crude model
```{r}
# Create two logistic models using glm() 
f0_exposure <- glm(HDL_clinic ~ PGS_HDL, data = mr_data, family = 'binomial')
f0_outcome <- glm(AD_bin ~ PGS_HDL, data = mr_data, family = 'binomial')
# Extract MR values
crude_mr <- get_MR_value(f0_exposure, f0_outcome, 'example_MR')
datatable(crude_mr)
```

Adjusted model
```{r}
# Create another two logistic models using glm() 
f1_exposure <- glm(HDL_clinic ~ PGS_HDL + age + sex + education, data = mr_data, family = 'binomial')
f1_outcome <- glm(AD_bin ~ PGS_HDL + age + sex + education, data = mr_data, family = 'binomial')
# Extract MR values
adj_mr <- get_MR_value(f1_exposure, f1_outcome, 'adj_MR')
# Bind results together
total_mr <- rbind(crude_mr, adj_mr)
datatable(total_mr)
```

&nbsp;
&nbsp;

# Table output
All the tables could be output to your local folder as Excel speadsheets, then you could simply do some formatting!
```{r eval=FALSE}
sheets_all <- list('logistic_table' = tab_OR_all,
                  'linear_table' = tab_linear_all,
                  'mr_table' = total_mr)
writexl::write_xlsx(sheets_all, path = paste0(output_path, 'table_in_one.xlsx'))
```

